---
title: LimeRick
output: github_document
---
 
<!-- README.md is generated from README.Rmd. Please edit this file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

<!--
`r badgecreatr::minimal_badges(status = "concept", license = "MIT", date = Sys.Date())`

```{r echo = FALSE}
#badgecreatr::badgeplacer(status = "wip" , githubaccount = "k127", githubrepo = "LimeRick")
#[![minimal R version](https://img.shields.io/badge/R>=-3.3.1-6666ff.svg)](https://cran.r-project.org/)
```

[![Project Status: active](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/LimeRick)](https://cran.r-project.org/package=LimeRick)
[![rstudio mirror downloads](http://cranlogs.r-pkg.org/badges/LimeRick?)](https://github.com/metacran/cranlogs.app)
-->

**Bridge between R and LimeSurvey**

This package is a _[LimeSurvey](http://limesurvey.org) RemoteControl 2_ JSON-RPC API Client for R.

## Motivation

* need for a bridge that **connects very closely** two open source projects (R&nbsp;and&nbsp;LimeSurvey)
* need for an **advanced tool** that allow for:
    * **quickly importing** responses into R from active online surveys
    * **automatically accessing** properties of survey and questions
    * **monitoring** survey status and responses directly from R
    * **adding new responses** to a survey directly from R
    * **simplifying workflow** for reproducible analysis <br/>
      (by providing in advance structured data schema)
    * developing **data products** based on **real-time** declarative data collection <br/>
      (e.g. continuous online tracking studies)
    * collecting meta-data of respondents interactions with online surveys <br/>
      on **unique low-granular level**

## Tutorial

### Installing from GitHub

```{r, echo=TRUE, eval=FALSE, cache=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("k127/LimeRick")
```

### Loading the installed package

```{r, echo=TRUE, eval=TRUE, cache=FALSE}
library(LimeRick)
```

### Configuring access to a survey

```{r, echo=TRUE, eval=TRUE, cache=FALSE}
# Set the LimeSurvey RemoteControl 2 JSON-RPC API URL
options(lsAPIurl = "https://survey.example.com/index.php/admin/remotecontrol")

# Set LimeSurvey credentials
options(lsUser = "LimeRickDemo")
options(lsPass = "LimeRickDemo")
```

### Low-level API function calls vs. function wrappers

```{r, echo=TRUE, eval=FALSE, cache=FALSE}
# Compare a low level API call …
lsAPI(method = "get_session_key", params = list(admin = getOption("lsUser"),
                                                password = getOption("lsPass")))

# … to this Higher level API call
lsGetSessionKey()
```

### Setting a connection to a survey

```{r, echo=TRUE, eval=FALSE, cache=FALSE}
# Get session key for the user and save it to the lsSessionCache environment
lsGetSessionKey()

# If you work with a specific locale at LimeSurvey you maybe want to set it
Sys.setlocale("LC_ALL", "Polish")
```

### Listing available surveys

```{r, echo=TRUE, eval=FALSE, cache=FALSE}
# List available surveys
(surveyList = lsListSurveys())

# Extract the ID of the first survey in the list
surveyID = surveyList$sid[1] 
```

### Listing survey questions

```{r, echo=TRUE, eval=FALSE, cache=FALSE}
questionList = lsListQuestions(surveyID, lang = "en")
```

### Accessing survey or question properties

We can access all available question and survey properties. For example:
```{r, echo=TRUE, eval=FALSE, cache=FALSE}
# Is the survey active? (Y - Yes)
lsGetSurveyProperties(surveyID)$active

# What is the main text of a given question?
lsGetQuestionProperties(13, lang = "en")$question

# Is the question mandatory? (Y - Yes)
lsGetQuestionProperties(13, lang = "en")$mandatory
```

### Checking survey response summary

```{r, echo=TRUE, eval=FALSE, cache=FALSE}
lsGetSummary(surveyID)
```

### Export responses

```{r, echo=TRUE, eval=FALSE, cache=FALSE}
d = lsExportResponses(surveyID, completionStatus = "complete")
tail(d)
```

### Adding responses to the survey

```{r, echo=TRUE, eval=FALSE, cache=FALSE}
# Show possible options for a particular question
ls_getAnswerOptions(questionID = 13, lang = "en")

# Create a response using question codes: surveyID+"X"+groupID+"X"+questionID
response = list('123456X1X12' = "LimeRick",
                '123456X1X17' = "Adding feedback directly from R",
                '123456X2X28' = "A2",# Academia
                '123456X2X29' = "174" # Poland
                )

# Add the above response to the survey
lsAddResponse(surveyID, response)
```

## See also

* Limesurvey GmbH. / LimeSurvey: An Open Source survey tool /LimeSurvey GmbH, Hamburg, Germany. URL http://www.limesurvey.org
