---
title: "Creating a catalogue of translatables"
output: rmarkdown::html_vignette
description: >
  Learn how pkgdown's automatic linking works, and how you can 
  customise it.
vignette: >
  %\VignetteIndexEntry{Creating a catalogue of translatables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Use case

* We want to create a catalogue (data frame)
* With all translatable texts of a survey and
* Ordered by `group_order` and `question_order`
* For groups it should contain
  * `group_name`
  * `description`
* For questions it should contain
  * `question`
  * `help`
  * Whether it is a sub-question (`parent_qid != 0`)
  * If it is a sub-question, `parent_qid` must be preserved in a way
* For list questions it should contain all answer options
  * `answerText`
  * `answerCode` must be preserved in a way
  * It must be connected to its question

The catalogue should then be sent to a translator who will return it with
replaced texts. Those texts should then be sent to the LimeSurvey application.

## Design

How could the catalogue look like?

| Text ID | Source text | Target text |
|---|---|---|
| `paste0()` | "A group title" | "Ein Gruppentitel" |
| … |  |  |

## Environment preparation

```{r}
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

if (!require("LimeRick")) {
  if (!require("devtools")) install.packages("devtools")
  devtools::install_github("k127/LimeRick")
}
library(LimeRick)
```


## Set up survey ID and source language, get a session key

```{r}
# The ID of the to be translated survey
surveyID = "259165"

# Let's say that our translation workflow has English as the source language
lang = "en"

# Get a session key for our future requests
lsGetSessionKey()
```


## Get the group list

```{r}
# Note that the field group order has to be numerical for the arrangement to work
groups <- arrange(lsListGroups(surveyID = surveyID), group_order)
```


## Get the question list

```{r}
questions <- arrange(lsListQuestions(surveyID = surveyID, lang = lang), question_order)

# Note that what we actually receive is not a list
class(questions)
```

# TODO suggest package "rapport"

## Iterate over groups, questions

```{r}
# Create an empty data frame
.key_text_id = "text_id"
.key_source_text = "source_text"
.key_target_text = "target_text"
.col_names = c(.key_text_id, .key_source_text, .key_target_text)
catalogue <- read.table(text = "",
                        colClasses = c("character", "character", "character"),
                        col.names = .col_names)

for (.i in 1:nrow(groups)) {
  .gid = groups[.i,]$gid
  
  # add group_name to catalogue
  # TODO if source_text exists, only append text_id
  .group_name =  groups[.i,]$group_name
  if (trimws(.group_name) != "") {
    .df_row_group_name = data.frame(paste0("GID__", .gid, "__group_name"), .group_name, "")
    names(.df_row_group_name) = .col_names
    catalogue = rbind(catalogue, .df_row_group_name)
  }

  # add description to catalogue
  # TODO if source_text exists, only append text_id
  .group_description = groups[.i,]$description
  if (trimws(.group_description) != "") {
    .df_row_group_description = data.frame(paste0("GID__", .gid, "__description"), .group_description, "")
    names(.df_row_group_description) = .col_names
    catalogue = rbind(catalogue, .df_row_group_description)
  }

  .group_questions = arrange(filter(questions, gid == .gid, parent_qid == 0), question_order)
  for (.k in 1:nrow(.group_questions)) {
    .qid = .group_questions[.k,]$qid
    .title = .group_questions[.k,]$title

    # catch question
    # TODO if source_text exists, only append text_id
    .question = .group_questions[.k,]$question
    if (trimws(.question) != "") {
      .df_row_questionn = data.frame(paste0(.title, "__question"), .question, "")
      names(.df_row_questionn) = .col_names
      catalogue = rbind(catalogue, .df_row_questionn)
    }

    # catch help
    # TODO if source_text exists, only append text_id
    .help = questions[.k,]$help
    if (!is.na(.help) && trimws(.help) != "") {
      .df_row_help = data.frame(paste0(.title, "__help"), .help, "")
      names(.df_row_help) = .col_names
      catalogue = rbind(catalogue, .df_row_help)
    }
    
    # are there sub-questions?
    .sub_questions = arrange(filter(questions, parent_qid == .qid), question_order)
    if (nrow(.sub_questions) > 0) for (.l in 1:nrow(.sub_questions)) {
      .sq_title = paste0(.title, "__", .sub_questions[.l,]$title)

      # catch sub-question
      .sub_question = .sub_questions[.l,]$question
      if (!is.na(.sub_question) && trimws(.sub_question) != "") {
        .df_row_sub_question = data.frame(paste0(.sq_title, "__question"), .sub_question, "")
        names(.df_row_sub_question) = .col_names
        catalogue = rbind(catalogue, .df_row_sub_question)
      }
    }
  # TODO answer options
  # TODO question validation tips and other translatables from question properties
  }
}
```


## Get answer options for each "L" type question

```{r}
# Type "L" seems to be the question type for lists, those own answer options
filter(questions, type == "L")[,"qid"]

ls_getAnswerOptions("170", lang = lang)
```







## Within a package

pkgdown will automatically link to documentation and articles wherever it's possible to do unambiguously. This includes:

* Bare function calls, like `build_site()`.
* Calls to `?`, like `?build_site` or `package?pkgdown`.
* Calls to `help()`, like `help("pkgdown")`.
* Calls to `vignette()`, like `vignette("pkgdown")`.

## Across packages

Linking to documentation in another package is straightforward. Just adapt the call in the usual way:

* `purrr::map()`, `MASS::addterm()`.
* `?purrr::map`, `?MASS::addterm`.
* `vignette("other-langs", package = "purrr")`, `vignette("longintro", package = "rpart")`

If pkgdown can find a pkgdown site for the remote package, it will link to it; otherwise it will link to <http://rdrr.io/> for documentation, and CRAN for vignettes. In order for a pkgdown site to be findable, it needs to be listed in two places:

*   In the `URL` field in the `DESCRIPTION`, as in
    [dplyr](https://github.com/tidyverse/dplyr/blob/85faf79c1fd74f4b4f95319e5be6a124a8075502/DESCRIPTION#L15):
  
    ```
    URL: http://dplyr.tidyverse.org, https://github.com/tidyverse/dplyr
    ```

*   In the `url` field in `_pkgdown.yml`, as in 
    [dplyr](https://github.com/tidyverse/dplyr/blob/master/_pkgdown.yml#L1)
    
    ```yaml
    url: https://dplyr.tidyverse.org
    ```
    
    When this field is defined, pkgdown generate a public facing
    [`pkgdown.yml` file](https://dplyr.tidyverse.org/pkgdown.yml) that provides
    metadata about the site:
    
    ```yaml
    pandoc: '2.2'
    pkgdown: 1.3.0
    pkgdown_sha: ~
    articles:
      compatibility: compatibility.html
      dplyr: dplyr.html
      dplyr_0.8.0: future/dplyr_0.8.0.html
      dplyr_0.8.0_new_hybrid: future/dplyr_0.8.0_new_hybrid.html
      programming: programming.html
      two-table: two-table.html
      window-functions: window-functions.html
    urls:
      reference: https://dplyr.tidyverse.org/reference
      article: https://dplyr.tidyverse.org/articles
    ```

Now, when you build a pkgdown site for a package that links to the dplyr documentation (e.g., `dplyr::mutate()`), pkgdown looks first in dplyr's `DESCRIPTION` to find its website, then it looks for `pkgdown.yml`, and uses the metadata to generate the correct links.
